FITS format for planetary surfaces: definitions, applications and best practices.

### 2018-03-01 ess_marmo.tex
Update on Overleaf.

\begin{keypoints}
\item We present an extension of FITS metadata description for planetary surface
investigations. 
\item A FITS description for planetary data aims
to simplify sharing data across the planetary and astronomy domains.
\item FITS open description metadata will promote interoperability from raw data
formatting to final visualization.
\end{keypoints}

@@

This paper proposes to use Flexible Image Transport
System\footnote{https://fits.gsfc.nasa.gov/}
(FITS) format \citep{fitsorig,fitsver3} in planetary surface investigations,
and describes how FITS metadata can easily be inserted in the
PDS4\footnote{https://pds.nasa.gov/tools/standards-reference.shtml}
metadata distribution model.
FITS is one of the standard formats implemented in the Virtual
Observatory\footnote{http://www.ivoa.net/astronomers/using\_the\_vo.html}
(VO), therefore making obvious its connection to the Planetary Virtual
Observatory initiative and the planetary Table Access Protocol (EPN-TAP)
\citep{ERARDepntap}.
FITS format is open, flexible, and largely implemented in open and efficient
processing tools allowing to handle large amounts of raster data.
The goal of this approach is to make easier data mining and reprocessing
in planetary surface investigations, promoting general software based on those
standards.
The option to use FITS within the planetary surface domain could help to
homogenize methods from raw formatting to visualization, while
allowing for optimized data processing across the planetary and astronomy domains.

@@

\subsubsection{FITS WCS keywords and their recommended use in planetary surface context}
\label{sssec:wcskeys}
In this section WCS FITS keywords are inserted in the planetary surface context.
When possible, corresponding PDS4 and EPN-TAP definitions are added for archiving
and distribution standardization purposes.
\begin{enumerate}
\item{RADESYS.
Name of the inertial reference frame with respect to which the coordinate
system is defined (defaulted to \texttt{`ICRS'})}
\item{EQUINOX.
The value field contains the equinox in years for the celestial
coordinate system in which positions are expressed (defaulted to \texttt{`J2000'})}
\item{WCSAXES.
Number of axes in the WCS description.
The value of WCSAXES may exceed the number of raster pixel axes. WCSAXES is, e.g.,
generally set to $2$ in two-dimensional rasters, it could be set to $3$ if an
additional time dependency is added to the coordinate description.}
\item{WCSNAME.
Name of world coordinate descriptions (e.g. body-fixed rotating -ocentric -ographic,
non-rotating, inertial).
EPN-TAP provides a \texttt{spatial\_coordinate\_description} parameter.}
\item{CRPIXn.
Location of a reference point along axis $n$, in units of the axis index.
The reference point value do not need to lie within the actual data array.}
\item{CTYPEn.
Name of the coordinate represented by axis $n$, in ``4--3'' form: the first four
characters specify the coordinate type, the fifth character is a `-' and the
remaining three characters specify an algorithm code for computing the world coordinate
value (\ref{table:proj}).
For angular coordinates \citet{calagreisenI} provides the form xLN xLG or yzLN yzLG,
where x or yz correspond to the target body.
We propose to standardize those codes as listed in Table \ref{table:ctype}.
For projected coordinates in meters we propose a similar scheme: xPX as Projected X,
xPY as Projected Y, or yzPX and yzPY.
Alternative representation of WCS are already possible in FITS, to describe images in
degrees and in meters \citep{greisencalaII}.
The name of the alternative axis will be specified using the alternative CNAMEn
keyword.}
\item{CRVALn.
Value of the coordinate specified by the CTYPEn keyword at the reference point CRPIXn.}
\item{CDi\_j.
The $ij$ linear transformation matrix (rotation, skewness, scaling) defining the
intermediate world coordinates.}
\item{PCi\_j.
The $ij$ linear transformation matrix (rotation, skewness) defining the intermediate
pixel coordinates (to be used with CDELT keywords).}
\item{CDELTn.
The coordinate increment along axis $n$ (scaling component of the linear transformation
matrix PCi\_j).
In PDS4 they correspond to the \textit{pixel\_resolution\_x} and
\texttt{pixel\_resolution\_y} attributes.
In EPN-TAP they are connected to the \texttt{spatial\_resolution} parameter.}
\item{CUNITn.
Units of the coordinate system, not mandatory, defaulted to degrees for angular
coordinates and to meters (m) for linear ones.}
\item{CNAMEn.
The description of a particular coordinate for an axis in a particular WCS (up to 68
characters), while the WCSNAMEa keyword names the particular WCS as a whole.
Its default value will be blank.}
\end{enumerate}

>>>>

\begin{keypoints}
\item We present an extension of FITS metadata for planetary surface
investigations. 
\item A FITS description for planetary data aims
to simplify sharing data across the astronomy and planetary domains.
\item An open FITS portrayal will promote interoperability from raw data
formatting to final visualization.
\end{keypoints}

@@

This paper proposes to use Flexible Image Transport
System\footnote{https://fits.gsfc.nasa.gov/}
(FITS) format \citep{fitsorig,fitsver3} in planetary surface investigations,
and describes how FITS metadata can easily be inserted in the
PDS4\footnote{https://pds.nasa.gov/tools/standards-reference.shtml}
metadata distribution model.
FITS is one of the standard formats implemented in the Virtual
Observatory\footnote{http://www.ivoa.net/astronomers/using\_the\_vo.html}
(VO), therefore making obvious its connection to the Planetary Virtual
Observatory initiative and the planetary Table Access Protocol (EPN-TAP)
\citep{ERARDepntap}.
FITS format is open, flexible, and largely implemented in open and efficient
processing tools allowing to handle large amounts of raster data.
The goal of this approach is to make easier data mining and reprocessing
in planetary surface investigations, promoting general software based on those
standards.
The option to use FITS within the planetary surface domain can help to
homogenize methods from raw formatting to visualization, while
allowing for optimized data processing across the planetary and astronomy domains.

@@

\subsubsection{FITS WCS keywords and their recommended use in planetary surface context}
\label{sssec:wcskeys}
In this section WCS FITS keywords are inserted in the planetary surface context.
When possible, corresponding PDS4 and EPN-TAP definitions are added for archiving
and distribution standardization purposes.
\begin{enumerate}
\item{RADESYS.
Name of the inertial reference frame with respect to which the coordinate
system is defined (defaulted to \texttt{`ICRS'})}
\item{EQUINOX.
The value field contains the equinox in years for the celestial
coordinate system in which positions are expressed (defaulted to \texttt{`J2000'})}
\item{WCSAXES.
Number of axes in the WCS description.
The value of WCSAXES may exceed the number of raster pixel axes. WCSAXES is, e.g.,
generally set to $2$ in two-dimensional rasters, it could be set to $3$ if an
additional time dependency is added to the coordinate description.}
\item{WCSNAME.
Name of world coordinate descriptions (e.g. body-fixed rotating -ocentric -ographic,
non-rotating, inertial).
EPN-TAP provides a \texttt{spatial\_coordinate\_description} parameter.}
\item{CRPIXn.
Location of a reference point along axis $n$, in units of the axis index.
The reference point value do not need to lie within the actual data array.}
\item{CTYPEn.
Name of the coordinate represented by axis $n$, in ``4--3'' form: the first four
characters specify the coordinate type, the fifth character is a `-' and the
remaining three characters specify an algorithm code for computing the world coordinate
value (\ref{table:proj}).
For angular coordinates \citet{calagreisenI} provides the form xLN xLT or yzLN yzLT,
where x or yz correspond to the target body.
We propose to standardize those codes as listed in Table \ref{table:ctype}.
For projected coordinates in meters we propose a similar scheme: xPX as Projected X,
xPY as Projected Y, or yzPX and yzPY.
Alternative representation of WCS are already possible in FITS, to describe images in
degrees and in meters \citep{greisencalaII}.
The name of the alternative axis will be specified using the alternative CNAMEn
keyword.}
\item{CRVALn.
Value of the coordinate specified by the CTYPEn keyword at the reference point CRPIXn.}
\item{CDi\_j.
The $ij$ linear transformation matrix (rotation, skewness, scaling) defining the
intermediate world coordinates.}
\item{PCi\_j.
The $ij$ linear transformation matrix (rotation, skewness) defining the intermediate
pixel coordinates (to be used with CDELT keywords).}
\item{CDELTn.
The coordinate increment along axis $n$ (scaling component of the linear transformation
matrix PCi\_j).
In PDS4 they correspond to the \textit{pixel\_resolution\_x} and
\texttt{pixel\_resolution\_y} attributes.
In EPN-TAP they are connected to the \texttt{spatial\_resolution} parameter.}
\item{CUNITn.
Units of the coordinate system, not mandatory, defaulted to degrees for angular
coordinates and to meters (m) for linear ones.}
\item{CNAMEn.
The description of a particular coordinate for an axis in a particular WCS (up to 68
characters), while the WCSNAMEa keyword names the particular WCS as a whole.
Its default value will be blank.}
\end{enumerate}
